Rule Synthesis for Rule-based Intrusion Detectors

### 2020-06-01 paper/paper.tex
saving space

Open-source \nids{}, like Snort~\cite{snort} and
Suricata~\cite{suricata}, provide rulesets for many existing
attacks. Unfortunately, attackers actively look for new mechanisms to
\begin{wrapfigure}[8]{r}{0.21\textwidth}
    \vspace{-2ex}  
    \centering
    \scalebox{0.9}{  
      \begin{tikzpicture}
        \begin{axis}[
            width = 0.3\textwidth,
            height = 0.2\textwidth,            
            bar width=3pt,
            scaled ticks=false,
            ymajorticks=false,
            ybar stacked,
            ymax=250,
            enlargelimits=0.09,
            clip=true,
            legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
            xtick=data,
            xticklabel style={rotate=45, font=\footnotesize},
            tick pos=left,
            nodes near coords,
            every node near coord/.append style={font=\footnotesize},
            nodes near coords align={vertical},
            symbolic x coords={Jan, Feb, Mar, Apr, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec},]
            \addplot coordinates {(Jan, 114) (Feb, 155) (Mar, 117) (Apr, 159) (May, 110) (Jun, 231) (Jul, 122) (Aug, 167) (Sep, 107) (Oct, 225) (Nov, 148) (Dec, 127)};
        \end{axis}
      \end{tikzpicture}
    }
    \vspace{-5ex}
    \caption{\label{fig:distribution-rules-per-month}2019 new rules.}
\end{wrapfigure}
exploit systems through the
network. Figure~\ref{fig:distribution-rules-per-month} shows the
number of rules added per month in the year of 2019 to the ``Emerging
Threats Open Ruleset''~\cite{emerging-threats-open}, a popular public
data set of \suri\ rules. A closer observation on the changelogs of
this ruleset~\cite{emerging-threats-changelogs} shows that new rules
are often created.  Rulesets need to be continuously revised as the
community finds and advertises vulnerabilities, which eventually can
be exploited by hackers.  Unfortunately, creating these rules is
time-consuming and error-prone~\cite{vollmer-etal-cics2011,alparslan-blog-suri}. The language for
describing rules is non-trivial (see
Section~\ref{sec:example-suricata-rules}) and testing correctness of a
rule requires data and careful inspection.

@@

\vspace{1ex}
\noindent\textbf{Overview.}~\tname\ takes on input one or more
examples of an attack and produces on output a ranked list of
\emph{plausible rules}, \ie{}, rules capturing only the malicious
traffic produced by the attacks. Figure~\ref{fig:overview} shows 
the workflow of \tname{} as a pipeline of three components, appearing 
numbered in the figure. As public data sets of benign
traffic~\cite{tcpreplay,stratosphere-normal} and rules
exist~\cite{emerging-threats-open}, a user of \tname{}
only needs to provide examples of attacks.

@@

\tname\ works as follows. First, it creates a rule that captures the
malicious traffic associated with example attacks provided on
input. In this step, \tname\ detects options directly from the input
messages. More precisely, it creates a rule containing options that
are guaranteed to be satisfied by the malicious traffic. We refer to
that rule as the \emph{seed rule} and its corresponding set of options
as \emph{seed options}. It is important to note that, because only the
malicious traffic is considered in this step, the seed rule is
potentially over-specified, \ie{}, it may be more restrictive than
necessary. Consequently, the use of that rule could result in false
negatives. Second, \tname\ uses benign traffic to generate alternative
rules that 1) weaken the constraints expressed in the seed rule while
2) still capturing the malicious traffic. Given that every option
declared in a rule needs to be satisfied for a rule to capture an
attack, any subset of the set of seed options, by definition, captures
the attack characterized by the malicious traffic. \tname{} leverages
that observation to search for plausible rules containing a proper
subset of options from the seed options that captures \emph{only} the
malicious traffic. Finally, \tname{} uses heuristic functions,
obtained from a data set of existing rules, to rank the generated
plausible rules by their similarity to existing rules. Similarity is
used to approximate user's intent. The following sections detail each
component of \tname.

@@

%\setlength{\tabcolsep}{5pt}
%renewcommand{\arraystretch}{0.9}

@@

\newcommand{\textRQone}{How well \tname\ ranks the golden rule?}
%\vspace{0.2cm}
\begin{itemize}[leftmargin=*,label={}]
\item{\textbf{RQ1.}} \textRQone
\end{itemize}

@@

\newcommand{\textRQtwo}{How important are the heuristic functions
  to \tname's performance?}
%\vspace{0.2cm}
\begin{itemize}[leftmargin=*,label={}]
\item{\textbf{RQ2.}} \textRQtwo\
\end{itemize}

@@

\newcommand{\textRQthree}{How sensitive \tname\ is to the amount of
  positive examples?} \rui{should we also investigate the impact of the negative ones?}
%\vspace{0.2cm}
\begin{itemize}[leftmargin=*,label={}]
\item{\textbf{RQ3.}} \textRQthree\
\end{itemize}

@@

\begin{table*}[t!]
  \footnotesize
  \setlength{\tabcolsep}{5pt}
  \renewcommand{\arraystretch}{-0.01}
  \caption{\label{table:comparison}Alternative techniques.}
  \vspace{-2ex}
  \centering
    \begin{tabular}{lcccccc}
    \toprule
    \multicolumn{1}{c}{Paper} &
    \multicolumn{1}{c}{Input} &
    \multicolumn{1}{c}{Output} &    
    \multicolumn{1}{c}{Dataset} &
    \multicolumn{1}{c}{Protocols} &
    \multicolumn{1}{c}{Evaluation} &
    \multicolumn{1}{c}{Avail.?} \\
    \midrule
    \cite{vollmer-etal-cics2011} & Malign Traffic & List  & Manually crafted & ICMP & TP, FP & No\\
%    \midrule
    \cite{fallahi} & Mixed Traffic & One Rule  & \cite{iscx} & HTTP, SSH, TCP & TP, FP, Precision, Recall, F1 & No\\    
%    \midrule
    \cite{Kao2015AutomaticNR} & Malign/Benign Traffic, Rule Dataset &
    One Rule & Manually crafted, \cite{emerging-threats-open} & HTTP & Generating time, FP & No \\
%    \midrule
    \cite{guruprasad}& Malign Traffic & List  & \cite{iscx} \cite{darpa1999} &
    ICMP, TCP & FP & No\\
%    \midrule
    Syrius & Malign/Benign Traffic, Rule Dataset & Ranking & Manually crafted, \cite{emerging-threats-open}, \cite{iscx} & HTTP, TCP, ICMP & Rule List & Yes \\
    \bottomrule
  \end{tabular}
\end{table*}

@@

\begin{figure}[h]
  \vspace{-2ex}
  \centering
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
    \caption{idqaccess}
  \end{subfigure}%
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}       \caption{scripttag}    
  \end{subfigure}%
  \caption{\label{fig:impact-number-attacks}Impact of number of positive examples.}
  \vspace{-2ex}
\end{figure}

Figure~\ref{fig:impact-number-attacks} shows the results for each
attack we selected. The y-axis of each plot denotes the rank of the
golden rule and x-axis denotes the number of attack examples used to
synthesize rules. Each point in the x-axis is associated with the
boxplot of the distribution for the corresponding number of
variants. Results indicate that the number of examples is important,
the choice of examples is important, and a small number of examples
suffices to reach the best results.

\begin{figure}[h!]
  \vspace{-1.5ex}
  \centering
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/coldfusion/coldfusion.pdf}
    \caption{coldfusion}
  \end{subfigure}%
  \hspace{-8ex}
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/inc/inc.pdf}
    \caption{inc}    
  \end{subfigure}%
  \hspace{-8ex}
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
    \caption{jsp}    
  \end{subfigure}%
  \caption{Impact of number of negative examples.}
  \label{fig:impact-negative}
  \vspace{-2ex}
\end{figure}

@@

\noindent
\textbf{Alternative methods for rule synthesis.} As detailed before,
we could not access the tools implementing the techniques for
comparison. Nevertheless, we found that there are important
differences between our work and prior work. Vollmer
\etal{}~\cite{vollmer-etal-cics2011} proposed the use of search to
derive rules for Snort, which are very similar to \suri's.  Their
solution focused on icmp attacks for which plausible attacks can be
created directly from packets (see
Section~\ref{sec:rules-and-packets}). They modeled every field of the
icmp packet in chromosomes of candidate solutions (\ie{}, a rule) of a
genetic search. The initial population is a set of candidate solutions
that randomly combine options, modelling the packet fields. They
showed that they were able to generate rules without false negatives
and only 3 false positives across 33K benign packets. The key
differences between our work and Vollmer~\etal{}'s are 1) the support
of different protocols and 2) the use of ranking to discriminate
alternative plausible rules. As for 1, our approach supports a variety
of protocols and options (\eg{}, options that refer to message
payloads). The icmp rules they generate only refer to message fields.
\MyComment{\Luc{eh possivel incluir contents em regras icmp tb. o q a
    secao 2.3 diz q \textbf{geralmente} as regras nao-http possuem
    apenas field options}. } As for 2, our solution additionally ranks
plausible options, which contain \emph{no} false positives or
negatives with respect to the benign and mallicious traffic
considered. Guruprasad \etal{}\cite{guruprasad} also proposes an
evolutionary framework, creating their own ICMP dataset as Vollmer,
but also using DARPA 1999 \cite{darpa1999} and ISCX 2012 \cite{iscx}
datasets for non-ICMP attacks. Although Guruprasad supports more
protocols than Vollmer, our approach still supports a larger set of
protocols and the ability to ranks rules. Muthuregunathan
\etal{}~\cite{muthuregunathan-etal2009} propose a clustering method to
create rules. They characterize candidate solutions as a set of field
values and use a clustering algorithm---hybrid k-means partitioned
around menoids---to isolate malicious from benign
traffic. Subsequently, they optimize the formed clusters using genetic
algorithms and hill-climbing search. Ou approach differs from theirs
in the method used, results, and support for content
options. Kao~\etal~\cite{Kao2015AutomaticNR} proposed ARGS, a system
to synthesize http rules that take malicious traffic as input. ARGS
looks for a common pattern across the different inputs to define
rules. It does not take existing rules as input or benign
traffic. Consequently, false positive can be produced and need to be
handled manually by users. Our approach, in contrast, does not
generate false positives w.r.t. the input traffic.

@@

\Gui{Fallahi \etal{}\cite{fallahi} proposes a flow-based solution for automated rule generation. Instead of analyzing the packets individually, their algorithm investigates and creates the rules for flows in the network. The algorithm was tested using the ISCX 2012 dataset and showed good results in terms of precision, but it is limited to the type of attack. As it won't analyse individual details of each packet, it wouldn't be able to identify and generate rules for attacks that require detailed options, such as the HTTP PESC, RCE and LFI.}

@@

\noindent
\textbf{Rule repair.}~Hallahan~\etal{}~\cite{8102263} propose to
repair firewall rules based on false positive detected by the
rules. Although firewall and \nids\ rules have differences and the
problem addressed is different---repair not synthesis---the work is
similar in spirit and inspiring. We remain to evaluate how \tname\ can
be used for repair.

\Gui{Ganesan \etal{}\cite{ganesan} introduces an idea of generating new sets of rules based on modifications of existing rules. They used the community edition of snort rules from 2012 for the modifications and tested with the the pcap file from the MACCDC 2012 dataset. It was reported an increase in the number of alerts, but it wasn't analysed the false alarmrates for abduced rules. Their technique is parallel to the rule creation strategies, as the objective of the modified rules is to capture variations and evolutions of attacks.}

\noindent
\textbf{Anomaly-based intrusion detection.}
Cordy~\etal{}\cite{cordy-etal-issta19} propose the use of a genetic
algorithm to not only generate attack schemes capable of penetrating
the tested anomaly-based NIDS, but also repairing the NIDS to identify
the given attack scheme. Although our approach and theirs use machine
intelligence to detect attacks, anomaly-based and rule-based
\nids\ are complementary, as discussed on
Section~\ref{sec:background}.

@@

Network Intrusion Detectors are an important defense mechanism against
networks attacks. This paper presents \tname{}, a novel approach to
generate rules for Rule-based \nids. \tname\ leverages positive and
negative traffic and examples of trusted rules to synthesize
rules. \tname{} is organized as a pipeline of three components that
1)~create an over-specified seed rule, 2)~derive plausible rules from
the seed, and 3)~rank plausible rules. Experimental results provide
early evidence showing that \tname\ is very promising. It was able to
rank the correct rule among the first five rules of the reported
ranking in \percTopFiveRanking\ of the cases we analyzed. The data set
containing our experimental artifacts is publicly accessible at
\ourdataset.

\rui{thinking out loud here: one of the motivations is that it is time-consuming
and troublesome to generate new rules. Time-consuming is not being addressed in 
the RQs and I think that, albeit it is difficult to conduct such a study, we 
shuold discuss it. I would think that mentinoing the human-on-the-loop (not in-the-loop)
should also be dscussed (this is more about the troublesomeness)}

>>>>

Open-source \nids{}, like Snort~\cite{snort} and
Suricata~\cite{suricata}, provide rulesets for many existing
attacks. Unfortunately, attackers actively look for new mechanisms to

\begin{wrapfigure}[8]{r}{0.21\textwidth}
    \vspace{-2ex}  
    \centering
    \scalebox{0.9}{  
      \begin{tikzpicture}
        \begin{axis}[
            width = 0.3\textwidth,
            height = 0.2\textwidth,            
            bar width=3pt,
            scaled ticks=false,
            ymajorticks=false,
            ybar stacked,
            ymax=250,
            enlargelimits=0.09,
            clip=true,
            legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
            xtick=data,
            xticklabel style={rotate=45, font=\footnotesize},
            tick pos=left,
            nodes near coords,
            every node near coord/.append style={font=\footnotesize},
            nodes near coords align={vertical},
            symbolic x coords={Jan, Feb, Mar, Apr, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec},]
            \addplot coordinates {(Jan, 114) (Feb, 155) (Mar, 117) (Apr, 159) (May, 110) (Jun, 231) (Jul, 122) (Aug, 167) (Sep, 107) (Oct, 225) (Nov, 148) (Dec, 127)};
        \end{axis}
      \end{tikzpicture}
    }
    \vspace{-5ex}
    \caption{\label{fig:distribution-rules-per-month}2019 new rules.}
\end{wrapfigure}
exploit systems through the
network. Figure~\ref{fig:distribution-rules-per-month} shows the
number of rules added per month in the year of 2019 to the ``Emerging
Threats Open Ruleset''~\cite{emerging-threats-open}, a popular public
data set of \suri\ rules. A closer observation on the changelogs of
this ruleset~\cite{emerging-threats-changelogs} shows that new rules
are often created.  Rulesets need to be continuously revised as the
community finds and advertises vulnerabilities, which eventually can
be exploited by hackers.  Unfortunately, creating these rules is
time-consuming and error-prone~\cite{vollmer-etal-cics2011,alparslan-blog-suri}. The language for
describing rules is non-trivial (see
Section~\ref{sec:example-suricata-rules}) and testing correctness of a
rule requires data and careful inspection.

@@

\vspace{1ex}
\noindent\textbf{Overview.}~\tname\ takes on input one or more
examples of an attack and produces on output a ranked list of
\emph{plausible rules}, \ie{}, rules capturing only the malicious
traffic produced by the attacks. Figure~\ref{fig:overview} shows the
workflow of \tname{} as a pipeline of three components, numbered in
the figure. As public data sets of benign
traffic~\cite{tcpreplay,stratosphere-normal} and rules
exist~\cite{emerging-threats-open}, \tname{}'s users only needs to
provide examples of attacks.

@@

\tname\ works as follows. First, it creates a rule that captures the
malicious traffic associated with example attacks provided on
input. In this step, \tname\ detects options directly from the input
messages. More precisely, it creates a rule containing options that
are guaranteed to be satisfied by the malicious traffic. We refer to
that rule as the \emph{seed rule} and its corresponding set of options
as \emph{seed options}. It is important to note that, because only the
malicious traffic is considered in this step, the seed rule is
potentially over-specified, \ie{}, it may be more restrictive than
necessary. Consequently, the use of that rule could result in false
negatives. Second, \tname\ uses benign traffic to generate alternative
rules that 1) weaken the constraints expressed in the seed rule while
2) still capturing the malicious traffic. Given that every option
declared in a rule needs to be satisfied for a rule to capture an
attack, any subset of the set of seed options, by definition, captures
the attack characterized by the malicious traffic. \tname{} leverages
that observation to search for plausible rules containing a proper
subset of options (from the seed options) that captures \emph{only} the
malicious traffic. Finally, \tname{} uses heuristic functions,
obtained from a data set of existing rules, to rank the generated
plausible rules by their similarity to existing rules. Similarity is
used to approximate user's intent. The following sections detail each
component of \tname.

@@

@@

\newcommand{\textRQone}{How well \tname\ ranks the golden rule?}
\begin{itemize}[leftmargin=*,label={}]
\item{\textbf{RQ1.}} \textRQone
\end{itemize}

@@

\newcommand{\textRQtwo}{How important are the heuristic functions
  to \tname's performance?}
%\vspace{0.2cm}

\begin{itemize}[leftmargin=*,label={}]
\item{\textbf{RQ2.}} \textRQtwo\
\end{itemize}

@@

\newcommand{\textRQthree}{How sensitive \tname\ is to the amount of
  positive and negative examples?}

\begin{itemize}[leftmargin=*,label={}]
\item{\textbf{RQ3.}} \textRQthree\
\end{itemize}

@@

\begin{table*}[t!]
  \footnotesize
  \setlength{\tabcolsep}{5pt}
  \renewcommand{\arraystretch}{-0.01}
  \caption{\label{table:comparison}Alternative techniques.}
  \vspace{-2ex}
  \centering
    \begin{tabular}{lcccccc}
    \toprule
    \multicolumn{1}{c}{Paper} &
    \multicolumn{1}{c}{Input} &
    \multicolumn{1}{c}{Output} &    
    \multicolumn{1}{c}{Dataset} &
    \multicolumn{1}{c}{Protocols} &
    \multicolumn{1}{c}{Evaluation} &
    \multicolumn{1}{c}{Avail.?} \\
    \midrule
    \cite{vollmer-etal-cics2011} & Malign Traffic & List  & Manually crafted & ICMP & TP, FP & No\\
%    \midrule
    \cite{fallahi} & Mixed Traffic & One Rule  & \cite{iscx} & HTTP, SSH, TCP & TP, FP, Precision, Recall, F1 & No\\    
%    \midrule
    \cite{Kao2015AutomaticNR} & Malign/Benign Traffic, Rule Dataset &
    One Rule & Manually crafted, \cite{emerging-threats-open} & HTTP & Generating time, FP & No \\
%    \midrule
    \cite{guruprasad}& Malign Traffic & List  & \cite{iscx} \cite{darpa1999} &
    ICMP, TCP & FP & No\\
%    \midrule
    Syrius & Malign/Benign Traffic, Rule Dataset & Ranking & Manually crafted, \cite{emerging-threats-open}, \cite{iscx} & HTTP, TCP, ICMP & Rule List & Yes \\
    \bottomrule
    \end{tabular}
    \vspace{-3ex}
\end{table*}

@@

\begin{wrapfigure}[11]{r}{0.28\textwidth}  
  \vspace{-2ex}
  \centering
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
    \vspace{-3ex}
    \caption{idqaccess}
  \end{subfigure}%
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
    \vspace{-3ex}        
    \caption{scripttag}
  \end{subfigure}%
  \caption{\label{fig:impact-number-attacks}Impact of number of positive examples.}
  \vspace{-2ex}
\end{wrapfigure}

Figure~\ref{fig:impact-number-attacks} shows results for each
attack we selected. The y-axis of each plot denotes the rank of the
golden rule and x-axis denotes the number of attack examples used to
synthesize rules. Each point in the x-axis is associated with the
boxplot of the distribution for the corresponding number of
variants. Results indicate that the number of examples is important,
the choice of examples is important, and a small number of examples
suffices to reach the best results.

\begin{figure}[h!]
  \vspace{-1.5ex}
  \centering
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/coldfusion/coldfusion.pdf}
    \vspace{-1ex}        
    \caption{coldfusion}
  \end{subfigure}%
  \hspace{-8ex}
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/inc/inc.pdf}
    \vspace{-1ex}    
    \caption{inc}    
  \end{subfigure}%
  \hspace{-8ex}
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
    \vspace{-1ex}    
    \caption{jsp}    
  \end{subfigure}%
  \caption{Impact of number of negative examples.}
  \label{fig:impact-negative}
  \vspace{-2ex}
\end{figure}

@@

\noindent
\textbf{Alternative methods for rule synthesis.} As detailed before,
we could not access the tools implementing the techniques for
comparison. Nevertheless, we found that there are important
differences between our work and prior work. Vollmer
\etal{}~\cite{vollmer-etal-cics2011} proposed the use of search to
derive rules for Snort, which are very similar to \suri's.  Their
solution focused on icmp attacks for which plausible attacks can be
created directly from packets (see
Section~\ref{sec:rules-and-packets}). They modeled every field of the
icmp packet in chromosomes of candidate solutions (\ie{}, a rule) of a
genetic search. The initial population is a set of candidate solutions
that randomly combine options, modelling the packet fields. They
showed that they were able to generate rules without false negatives
and only 3 false positives across 33K benign packets. The key
differences between our work and Vollmer~\etal{}'s are 1) the support
of different protocols and 2) the use of ranking to discriminate
alternative plausible rules. As for 1, our approach supports a variety
of protocols and options (\eg{}, options that refer to message
payloads). The icmp rules they generate only refer to message fields.
\MyComment{\Luc{eh possivel incluir contents em regras icmp tb. o q a
    secao 2.3 diz q \textbf{geralmente} as regras nao-http possuem
    apenas field options}. } As for 2, our solution additionally ranks
plausible options, which contain \emph{no} false positives or
negatives with respect to the benign and mallicious traffic
considered. Guruprasad \etal{}\cite{guruprasad} also proposes an
evolutionary framework, creating their own ICMP dataset as Vollmer,
but also using DARPA 1999 \cite{darpa1999} and ISCX 2012 \cite{iscx}
datasets for non-ICMP attacks. Although Guruprasad supports more
protocols than Vollmer, our approach still supports a larger set of
protocols and the ability to ranks rules. Muthuregunathan
\etal{}~\cite{muthuregunathan-etal2009} propose a clustering method to
create rules. They characterize candidate solutions as a set of field
values and use a clustering algorithm---hybrid k-means partitioned
around menoids---to isolate malicious from benign
traffic. Subsequently, they optimize the formed clusters using genetic
algorithms and hill-climbing search. Ou approach differs from theirs
in the method used, results, and support for content
options. Kao~\etal~\cite{Kao2015AutomaticNR} proposed ARGS, a system
to synthesize http rules that take malicious traffic as input. ARGS
looks for a common pattern across the different inputs to define
rules. It does not take existing rules as input or benign
traffic. Consequently, false positive can be produced and need to be
handled manually by users. Our approach, in contrast, does not
generate false positives w.r.t. the input traffic. Fallahi
\etal{}\cite{fallahi} proposed flow-based synthesis approach to rule
generation. Instead of analyzing packets individually, their algorithm
investigates and creates the rules for flows in the network. The
algorithm was tested using the ISCX 2012 dataset and showed good
results in terms of precision, but it is limited to that type of
attack. As it won't analyse individual details of each packet, it
wouldn't be able to identify and generate rules for attacks that
require detailed options, such as the HTTP PESC, RCE and LFI.
\textbf{Rule repair.}~Hallahan~\etal{}~\cite{8102263} propose to
repair firewall rules based on false positive detected by the
rules. Although firewall and \nids\ rules have differences and the
problem addressed is different---repair not synthesis---the work is
similar in spirit and inspiring. We remain to evaluate how \tname\ can
be used for repair.  Ganesan \etal{}\cite{ganesan} applied similar
ideas in the context of IDS, \ie{}, they proposed to adjust existing
rules with the goal of capturing variations of existing attacks. They
used the community edition of snort rules from 2012 for the
modifications and tested them with pcap file from the MACCDC 2012
dataset. Their technique is complementary to \tname{}.
\textbf{Anomaly-based intrusion detection.}
Cordy~\etal{}\cite{cordy-etal-issta19} used genetic search to not only
to generate attacks capable of penetrating the tested anomaly-based
NIDS, but also to repair the NIDS to identify these
attacks. \tname\ and their approach are complementary as anomaly-based
and rule-based \nids\ are complementary (see
Section~\ref{sec:background}).

%% It was reported an increase in the number of alerts, but it
%% wasn't analysed the false alarmrates for abduced rules.
%% parallel to the rule creation strategies, as the
%% objective of the modified rules is to capture variations and
%% evolutions of attacks.

\noindent

@@

@@

@@

Network Intrusion Detectors are an important defense mechanism against
networks attacks. Generating rules to capture new attacks can be
time-consuming and error-prone. This paper presents \tname{}, a novel
approach to generate rules for Rule-based \nids. \tname\ leverages
positive and negative traffic and examples of trusted rules to
synthesize rules. \tname{} is organized as a pipeline of three
components that 1)~create an over-specified seed rule, 2)~derive
plausible rules from the seed, and 3)~rank plausible
rules. Experimental results provide early evidence showing that
\tname\ is promising. It was able to rank the correct rule among the
first five rules of the reported ranking in \percTopFiveRanking\ of
the cases we analyzed, for example. The data set containing our
experimental artifacts is publicly accessible at \ourdataset.

%\rui{thinking out loud here: one of the motivations is that it is
%time-consuming and troublesome to generate new rules. Time-consuming
%is not being addressed in the RQs and I think that, albeit it is
%difficult to conduct such a study, we shuold discuss it. I would
%think that mentinoing the human-on-the-loop (not in-the-loop) should also be dscussed (this is more about the troublesomeness)}
